Can Symmetric Cryptography Be Liberated from the Von Neumann Style?

### 2024-10-12 paper.tex
Grammar

The present work aims to cast these two notions in the language of
functional programming. We shall see that bitslicing corresponds to a
run-off-the-mill \emph{data} representation change (in effect, a
matrix transposition). But this is only half of the work: our ability
to systematically transform the \emph{code} operating on such
representation relies on suitably polymorphic definitions:
parametricity at its best! Fixslicing, on the other hand, will be
rationalized through equational reasoning over purely functional
terms. Identifying commuting terms will be of key importance there:
programming with algebraic structures (rather than untamed arrays of
bits) will be instrumental!

>>>>

The present work aims at casting these two notions in the language of
functional programming. We shall see that bitslicing corresponds to a
run-off-the-mill \emph{data} representation change (in effect, a
matrix transposition). But this is only half of the work: our ability
to systematically transform the \emph{code} operating on such
representation rests on suitably polymorphic definitions:
parametricity at its best! Fixslicing, on the other hand, will be
rationalized through equational reasoning over purely functional
terms. Identifying commuting terms will be of key importance there:
programming with algebraic structures (rather than untamed arrays of
bits) will be instrumental!
